---
title: "Paper Methods and Results"
author: "Diego Montecino-Latorre"
date: "3/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r packages, include =T, message=FALSE, warning=FALSE}
library(rnaturalearthdata)
library(rnaturalearth)
library(ggmap)
library(ggspatial)
library(sf)
library(dplyr)
library(maps)
library(cowplot)
library(magick)
library(egg)
library(rstan)
library(scales) 
library(rethinking)
library(tidybayes)
library(reshape2)
library(lvplot)
library(RColorBrewer)
library(gridExtra)
library(ggplot2)
library(loo)
library(DescTools)
library(openxlsx)
library(grid)
library(kableExtra)
library(XML)
library(methods)
library(rentrez)
```

# METHODS

Fig 1. Location of the studied Eidolon helvum roosts. Panel A shows the position of the roost in in Africa. Panel B shows some of the trees occupied at the 37 Military Hospital and Panel C at the Kikundi Market.

```{r fig 1 ,message=FALSE, warning=FALSE}

#PLot
 main.plot= ggplot() + 
  geom_sf(data= africa, fill="white", color="white", lwd=0.2) +
  geom_sf(data = tanzania.ghana, fill="grey80", color="grey70", lwd=0.2) +
  geom_sf(data = accra.morogoro, size = 7, shape = 21, color=alpha("tomato", 0), fill=alpha("red", 0.9), inherit.aes = FALSE) +
  geom_sf_text(data = accra.morogoro, 
               aes(label=name), hjust="center", vjust=1.8, nudge_x = 1.5, size=10) +
  annotation_scale(location = "bl", width_hint = 0.6, line_width = 1, text_cex = 2, height= unit(0.5, "cm")) +
  theme(panel.background = element_rect(fill = 'light blue', colour = NA),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        plot.margin=margin(0,0,0,0, "cm"),
        panel.border=element_blank(), 
        panel.spacing = unit(0, "cm"),
        axis.line=element_blank()) +
  annotate(geom="text", x=50.2, y=36, label="A", color="black", size=9)
   
  
# morogoro and accra colonies
 
 # Example with PNG (for fun, the OP's avatar - I love the raccoon)
 accra.picture=ggdraw() +
   draw_image("accra.jpeg") + 
   annotate(geom="text", x=0.05, y=0.82, label="B", color="black", size=9)
 
 # draw_plot(my_plot)
 morogoro.picture=ggdraw() +
   draw_image("Morogoro.png") +
   annotate(geom="text", x=0.05, y=0.82, label="C", color="black", size=9)
 
 

 figure1=
 grid.arrange(
   grobs = list(accra.picture, morogoro.picture, main.plot),
   widths = c(0.8,1.3),
   heights = c(1, 1),
   layout_matrix = rbind(c(1, 3),
                         c(2, 3))
 )

figure1 #test 3

#ggsave(file="Fig1.tiff", figure1)



```


### Assessing seasonality

The following R code correspond to the STAN script for the intercept-only model.

```{r intercept model,message=FALSE, warning=FALSE}

modelstring.int="
data{
int N; // number of observations. It means it is an integer with nodecimal points
int y[N]; //outcome 
}

 parameters {
   // Define parameters to estimate
   real alpha;
 }   

model {
  real mu[N];
  
  // Priors
  alpha ~ normal(0, 1);
 
  // model
  for (i in 1:N) {
    mu[i] = alpha;
  }


  // Likelihood part of Bayesian inference
  for (i in 1:N) {
    y[i] ~ bernoulli_logit(mu[i]);
  }
}

generated quantities {
int y_rep[N];
vector[N] log_lik; 
real dev;
dev = 0;


for (i in 1:N){ 

real mu;
mu = alpha;

dev = dev + (-2)*bernoulli_logit_lpmf( y[i] | mu);
log_lik[i] = bernoulli_logit_lpmf(y[i] | mu);

y_rep[i] = bernoulli_logit_rng(alpha);

}

}
"
```

The following R code correspond to the STAN script for the sine - cosine model. 

```{r sine cosine model,message=FALSE, warning=FALSE}

modelstring.sine.cosine="
data{
int N; // number of observations. It means it is an integer with nodecimal points
int y[N]; //outcome 
real day_over_year[N];
}

 parameters {
   // Define parameters to estimate
   real alpha;
   real beta_sine;
   real beta_cosine;
 }   

model {
  real mu[N];
  
  // Priors
  alpha ~ normal(0, 1);
  beta_sine ~ normal(0, 1);
  beta_cosine ~ normal(0, 1);
 
  // model
  for (i in 1:N) {
    mu[i] = alpha + beta_cosine* cos(2*pi()*day_over_year[i]) + beta_sine* sin(2*pi()*day_over_year[i]);
  }


  // Likelihood part of Bayesian inference
  for (i in 1:N) {
    y[i] ~ bernoulli_logit(mu[i]);
  }
}

generated quantities {
int y_rep[N];
vector[N] log_lik; 
real dev;
dev = 0;


for (i in 1:N){ 

real mu;
mu = alpha + beta_cosine* cos(2*pi()*day_over_year[i]) + beta_sine* sin(2*pi()*day_over_year[i]);

dev = dev + (-2)*bernoulli_logit_lpmf( y[i] | mu);
log_lik[i] = bernoulli_logit_lpmf(y[i] | mu);

y_rep[i] = bernoulli_logit_rng(alpha + beta_cosine* cos(2*pi()*day_over_year[i]) + beta_sine* sin(2*pi()*day_over_year[i]));

}

}
"
```

Load data

```{r load data plot,,message=FALSE, warning=FALSE}

tz.dat=read.csv("data_morogoro_roost_final.csv")
gh.dat=read.csv("data_accra_roost_final.csv")

```

### Run the models

The intercept only model

```{r run intercept only model, message=FALSE, warning=FALSE, eval=FALSE}

# Data
N<-nrow(tz.dat)
y<-tz.dat$pos

data=list(N=N, y=y)

# Intercpt Only Model

mod_intercept_stan <- stan(model_code = modelstring.int,
                  data = data, control = list(adapt_delta=0.999, max_treedepth =17),
                  cores=4, chains=4, warmup = 5000, iter = 10000, thin = 5)

saveRDS(mod_intercept_stan, "model_intercept.RDS")


```


The sine-cosine only model

```{r run sine-cosine model, message=FALSE, warning=FALSE, eval=FALSE}


N<-nrow(tz.dat)
y<-tz.dat$pos
day_over_year<-tz.dat$date/365


data=list(N=N,
          y=y,
          day_over_year=day_over_year)


mod_sine_cos_stan <- stan(model_code = modelstring.sine.cosine,
                  data = data, control = list(adapt_delta=0.999, max_treedepth =17),
                  cores=4, chains=4, warmup = 5000, iter = 10000, thin = 5)

saveRDS(mod_sine_cos_stan, "model_sine_cosine.RDS")

```


### Assessing the association between reproductive season and coronavirus shedding

The following R code correspond to the STAN script for the fixed effects model.

```{r  fixed effects model, message=FALSE, warning=FALSE}

# -----  Fixed effects model ----#

modelstring_fixed=
  "
data {
  int N;
  int y[N];
  int lact[N];
  int wean[N];
}

parameters {
  real alpha;
  real beta_wean;
  real beta_lact;
  
}

model {
   
   for (i in 1:N){
      
      y[i] ~ bernoulli_logit(alpha + beta_wean*wean[i] + beta_lact*lact[i]);}
  
  alpha ~ normal( 0.0, 1);
  beta_wean ~ normal( 0.0, 1);
  beta_lact ~ normal( 0.0, 1);
}

generated quantities {
int y_rep[N];

for (i in 1:N){
y_rep[i] = bernoulli_logit_rng(alpha + beta_wean*wean[i] + beta_lact*lact[i]);}

}

"

```

The following R code correspond to the STAN script for the hiearchical model.

```{r  hierarchical model, message=FALSE, warning=FALSE}

#----- Grouping factor per month and the months included ----- #
# in reproductive period ------------------------------------- #

modelstring_hierachical=
  "
data {
  // Number of level-1 observations (an integer)
  int<lower=0> N;
  // Number of month clusters
  int<lower=0> J;
  // Number of repro clusters
  int<lower=0> K;

  //Outcome
  int y[N];

  // Cluster indexes
  int<lower=1> index_repro_periods[J];
  int<lower=1> index_months[N];
  
}

parameters {
  // Define parameters to estimate
  
  // Level-2 random effect
  real beta_month[J];
  real<lower=0> sigma_month;

  // Level-3 random effect
  real beta_repro[K];
  real<lower=0> sigma_repro;
}

transformed parameters  {
  // Varying intercepts
  real beta_0jk[J];
  real beta_0k[K];

  // Individual mean
  real p[N];

  // Varying intercepts definition
  
  // Level-3 (10 level-3 random intercepts)
  for (k in 1:K) {
  beta_0k[k] =beta_repro[k]* sigma_repro;
  }
  
  // Level-2 (100 level-2 random intercepts)
  for (j in 1:J) {
    beta_0jk[j] = beta_0k[index_repro_periods[j]] + beta_month[j] *sigma_month;
  }
  // Individual mean
  for (i in 1:N) {
    p[i] =beta_0jk[index_months[i]];
  }
}

model {
  // Prior part of Bayesian inference

  // Random effects distribution
  
  beta_month ~ normal(0, 1);
  beta_repro ~ normal(0, 1);
  

  sigma_month ~ cauchy(0,1) ;
  sigma_repro ~ cauchy(0,1) ;

  // Likelihood part of Bayesian inference
  for (i in 1:N) {
    y[i] ~ bernoulli_logit(p[i]);
  }
}


generated quantities {

int y_rep[N];
vector[N] log_lik; 
real dev;
dev = 0;


for (i in 1:N){ 

real mu;
mu = beta_0jk[index_months[i]];

dev = dev + (-2)*bernoulli_logit_lpmf( y[i] | mu);
log_lik[i] = bernoulli_logit_lpmf(y[i] | mu);

y_rep[i] = bernoulli_logit_rng(p[i]);

}

}
"
```

### Run the models 

Run the fixed effects model

```{r run fix effects model, message=FALSE, warning=FALSE, eval=FALSE}

N<-nrow(tz.dat)
y<-tz.dat$positive
wean=ifelse(tz.dat$repro.per=="weaning",1,0)
lact=ifelse(tz.dat$repro.per=="birth",1,0)

data=list(N=N, 
          y=y,
          wean=wean,
          lact=lact)


mod_fixed_stan <- stan(model_code = modelstring_fixed,  
                  data = data, control = list(adapt_delta=0.999, max_treedepth =17),
                  cores=4, chains=4, warmup = 5000, iter = 10000, thin = 5)

# saveRDS(mod_fixed_stan, "model_repro_periods_only_fixed_effects.RDS")  




```

Run the hierarchical model

```{r run hierarchical model, message=FALSE, warning=FALSE, eval=FALSE}

N <- nrow(tz.dat)
K <- length(unique(tz.dat$repro))
J <- length(unique(tz.dat$month))
y <- tz.dat$pos

index_months <- as.integer(factor(tz.dat$month, levels = month.name[c(8:12,1:7)]))

index_repro_periods=sapply(month.name[c(8:12,1:7)], function(x)
  as.integer(factor(tz.dat$repro, levels = c("rest", "birth", "weaning")))[tz.dat$month==x][1])


data=list(N=N, 
          K=K,
          J=J,
          y=y,
          index_repro_periods=index_repro_periods, 
          index_months=index_months)


mod_hierarchical_stan <- stan(model_code = modelstring_hierachical,  
                      data = data, control = list(adapt_delta=0.999, max_treedepth =17),
                      cores=4, chains=4, warmup = 5000, iter = 10000, thin = 5)


# saveRDS(mod_hierarchical_stan, "hierarchical_model.RDS")  

```

# RESULTS


Fig 2. Eidolon helvum monthly abundance (green line), precipitation (blue line), and coronavirus shedding (red line) at the roost in Morogoro, Tanzania (left); and at the roost in Accra, Ghana (right). Color bands indicate the “lactation” (orange), “weaning” (purple), and “rest of the year” (blue) reproductive periods.

```{r fig 2, message=FALSE, warning=FALSE}


# periods, precipitaiton and roost size

dat.acc=data.frame(
  
  # Accra
  pop=tapply(gh.dat$pop, list(gh.dat$month), unique)[match(month.name[c(3:12,1,2)], names(tapply(gh.dat$pop, list(gh.dat$month), unique)))],
  
  # Monthly Precipitaiton
  precip=tapply(gh.dat$precip.month, list(gh.dat$month), unique)[match(month.name[c(3:12,1,2)], names(tapply(gh.dat$precip.month, list(gh.dat$month), unique)))],
  
  #Positivity
  pos=tapply(gh.dat$positive, list(gh.dat$month), mean)[match(month.name[c(3:12,1,2)], names(tapply(gh.dat$positive, list(gh.dat$month), mean)))]+0.0022)




#MOROGORO

dat.mor=data.frame(
  # Population
  pop=scale(tapply(tz.dat$pop, list(tz.dat$month), unique)[match(month.name[c(8:12,1:7)], names(tapply(tz.dat$pop, list(tz.dat$month), unique)))], center = T)[,1],
  
  # Monthly Precipitaiton
  precip=scale(tapply(tz.dat$precip.month, list(tz.dat$month), unique)[match(month.name[c(8:12,1:7)], names(tapply(tz.dat$precip.month, list(tz.dat$month), unique)))], center = T)[,1],
  
  #Positivity
  pos=scale(tapply(tz.dat$positive, list(tz.dat$month), mean)[match(month.name[c(8:12,1:7)], names(tapply(tz.dat$precip.month, list(tz.dat$month), unique)))], center = T)[,1])





test=scale(tapply(tz.dat$pop, list(tz.dat$month), unique)[match(month.name[c(8:12,1:7)], names(tapply(tz.dat$pop, list(tz.dat$month), unique)))], center = T)

test2=scale(tapply(tz.dat$precip.month, list(tz.dat$month), unique)[match(month.name[c(8:12,1:7)], names(tapply(tz.dat$precip.month, list(tz.dat$month), unique)))], center = T)

test3=scale(tapply(tz.dat$positive, list(tz.dat$month), mean)[match(month.name[c(8:12,1:7)], names(tapply(tz.dat$precip.month, list(tz.dat$month), unique)))], center = T)


grob <- grobTree(textGrob("Rest of the year", x=0.1,  y=0.85, hjust=0,
                          gp=gpar(col="black", fontsize=13)))

grob2 <- grobTree(textGrob("Lactation", x=0.39,  y=0.85, hjust=0,
                           gp=gpar(col="black", fontsize=13)))

grob3 <- grobTree(textGrob("Weaning", x=0.64,  y=0.85, hjust=0,
                           gp=gpar(col="black", fontsize=13)))


# precipitation

precip.mor=ggplot() + 
  
  geom_polygon(aes(x=c(1, 10, 10,1), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(22, 25, 25,22), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(10, 14, 14,10), y = c(-1.1,-1.1,3,3)), fill=alpha("#FEE0B6",0.5))+
  geom_polygon(aes(x=c(14, 22, 22, 14), y = c(-1.1,-1.1,3,3)), fill=alpha("#DECBE4",0.5))+
  
  geom_line(data = dat.mor, aes(x = seq(2,24, by = 2), y = precip), color = alpha('darkgreen', 0.7)) +
  geom_point(data = dat.mor, aes(x = seq(2,24, by = 2,), y = precip, size=2), color = alpha('darkgreen', 0.7)) +
  
  scale_y_continuous(name ="Precipitation (mm/day)",
                     breaks=(seq(0,4, length.out = 6)-attr(test2, 'scaled:center')) / attr(test2, 'scaled:scale'),
                     labels=seq(0,4, length.out = 6)) +
  
  scale_x_continuous(name = "",
                     breaks=seq(2,25, by = 2),
                     labels=month.name[c(8:12, 1:7)]) +
  
  annotation_custom(grob) +
  annotation_custom(grob2) +
  annotation_custom(grob3) +
  
  theme(legend.position = "none", 
        axis.text.x = element_text(size=14, angle=90), 
        axis.text.y = element_text(size=14),
        axis.title.y = element_text(size=24, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=24, margin = margin(t = 15, r = 0, b = 0, l = 0)))


# precip.mor


# population

pop.mor=ggplot() + 
  
  geom_polygon(aes(x=c(1, 10, 10,1), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(22, 25, 25,22), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(10, 14, 14,10), y = c(-1.1,-1.1,3,3)), fill=alpha("#FEE0B6",0.5))+
  geom_polygon(aes(x=c(14, 22, 22, 14), y = c(-1.1,-1.1,3,3)), fill=alpha("#DECBE4",0.5))+
  
  geom_line(data = dat.mor, aes(x = seq(2,24, by = 2), y = pop), color = alpha('darkblue', 0.7)) +
  geom_point(data = dat.mor, aes(x = seq(2,24, by = 2,), y = pop, size=2), color = alpha('darkblue', 0.7)) +
  
  scale_y_continuous(name =expression(paste('Population size ', "(*10"^"4", ")")),
                     breaks=(seq(from = 0, to = 50000, length.out = 6)-attr(test, 'scaled:center')) / attr(test, 'scaled:scale'),
                     labels=seq(from = 0, to = 5, length.out = 6)) +
  
  scale_x_continuous(name = "", 
                     breaks=seq(2,25, by = 2),
                     labels=month.name[c(8:12, 1:7)]) +
  
  theme(legend.position = "none", 
        axis.text.x = element_text(size=14, angle=90), 
        axis.text.y = element_text(size=14),
        axis.title.y = element_text(size=24, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=24, margin = margin(t = 15, r = 0, b = 0, l = 0)))


# pop.mor


# positives

pos.mor=ggplot() + 
  
  geom_polygon(aes(x=c(1, 10, 10,1), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(22, 25, 25,22), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(10, 14, 14,10), y = c(-1.1,-1.1,3,3)), fill=alpha("#FEE0B6",0.5))+
  geom_polygon(aes(x=c(14, 22, 22, 14), y = c(-1.1,-1.1,3,3)), fill=alpha("#DECBE4",0.5))+
  
  geom_line(data = dat.mor, aes(x = seq(2,24, by = 2), y = pos), color = 'red') +
  geom_point(data = dat.mor, aes(x = seq(2,24, by = 2,), y = pos, size=2), color = "red") +
  
  scale_y_continuous(name ="Proportion positive to CoV",
                     breaks=(seq(0,0.3, by = 0.1)-attr(test3, 'scaled:center')) / attr(test3, 'scaled:scale'),
                     labels=seq(0,0.3, by = 0.1)) +
  
  scale_x_continuous(name = "Month",
                     breaks=seq(2,25, by = 2),
                     labels=month.name[c(8:12, 1:7)]) +
  
  theme(legend.position = "none", 
        axis.text.x = element_text(size=14, angle=90), 
        axis.text.y = element_text(size=14),
        axis.title.y = element_text(size=24, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=24, margin = margin(t = 15, r = 0, b = 0, l = 0)))


# pos.mor




## GHANA ##

test=scale(tapply(gh.dat$pop, list(gh.dat$month), unique)[match(month.name[c(3:12,1,2)], names(tapply(gh.dat$pop, list(gh.dat$month), unique)))], center = T)

test2=scale(tapply(gh.dat$precip.month, list(gh.dat$month), unique)[match(month.name[c(3:12,1,2)], names(tapply(gh.dat$precip.month, list(gh.dat$month), unique)))], center = T)

test3=scale(tapply(gh.dat$positive, list(gh.dat$month), mean)[match(month.name[c(3:12,1,2)], names(tapply(gh.dat$precip.month, list(gh.dat$month), unique)))], center = T)


grob <- grobTree(textGrob("Lactation", x=0.16,  y=0.85, hjust=0,
                          gp=gpar(col="black", fontsize=13)))

grob2 <- grobTree(textGrob("Weaning", x=0.4,  y=0.85, hjust=0,
                           gp=gpar(col="black", fontsize=13)))

grob3 <- grobTree(textGrob("Rest of the year", x=0.67,  y=0.85, hjust=0,
                           gp=gpar(col="black", fontsize=13)))


dat.acc=data.frame(
  # Population
  pop=scale(tapply(gh.dat$pop, list(gh.dat$month), unique)[match(month.name[c(3:12,1,2)], names(tapply(gh.dat$pop, list(gh.dat$month), unique)))], center = T)[,1],
  
  # Monthly Precipitaiton
  precip=scale(tapply(gh.dat$precip.month, list(gh.dat$month), unique)[match(month.name[c(3:12,1,2)], names(tapply(gh.dat$precip.month, list(gh.dat$month), unique)))], center = T)[,1],
  
  #Positivity
  pos=scale(tapply(gh.dat$positive, list(gh.dat$month), mean)[match(month.name[c(3:12,1,2)], names(tapply(gh.dat$precip.month, list(gh.dat$month), unique)))], center = T)[,1])


# precipitation

precip.acc=ggplot() + 
  
  geom_polygon(aes(x=c(1, 4, 4,1), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(16, 25, 25,16), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(4, 8, 8, 4), y = c(-1.1,-1.1,3,3)), fill=alpha("#FEE0B6",0.5))+
  geom_polygon(aes(x=c(8, 16, 16, 8), y = c(-1.1,-1.1,3,3)), fill=alpha("#DECBE4",0.5))+
  
  geom_line(data = dat.acc, aes(x = seq(2,24, by = 2), y = precip), color = alpha('darkgreen', 0.7)) +
  geom_point(data = dat.acc, aes(x = seq(2,24, by = 2,), y = precip, size=2), color = alpha('darkgreen', 0.7)) +
  
  scale_y_continuous(name ="",
                     breaks=(seq(0,6, length.out = 6)-attr(test2, 'scaled:center')) / attr(test2, 'scaled:scale'),
                     labels=seq(0,6, length.out = 6)) +
  
  scale_x_continuous(name = "",
                     breaks=seq(2,25, by = 2),
                     labels=month.name[c(3:12,1,2)]) +
  
  annotation_custom(grob) +
  annotation_custom(grob2) +
  annotation_custom(grob3) +
  
  theme(legend.position = "none", 
        axis.text.x = element_text(size=14, angle=90), 
        axis.text.y = element_text(size=14),
        axis.title.y = element_text(size=24, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=24, margin = margin(t = 15, r = 0, b = 0, l = 0)))


# precip.acc


# population

pop.acc=ggplot() + 
  
  geom_polygon(aes(x=c(1, 4, 4,1), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(16, 25, 25,16), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(4, 8, 8, 4), y = c(-1.1,-1.1,3,3)), fill=alpha("#FEE0B6",0.5))+
  geom_polygon(aes(x=c(8, 16, 16, 8), y = c(-1.1,-1.1,3,3)), fill=alpha("#DECBE4",0.5))+
  
  geom_line(data = dat.acc, aes(x = seq(2,24, by = 2), y = pop), color = alpha('darkblue', 0.7)) +
  geom_point(data = dat.acc, aes(x = seq(2,24, by = 2,), y = pop, size=2), color = alpha('darkblue', 0.7)) +
  
  scale_y_continuous(
                     name =' ',
                     breaks=(seq(from = 0, to = 1200000, length.out = 6)-attr(test, 'scaled:center')) / attr(test, 'scaled:scale'),
                     labels=seq(from = 0, to = 120, length.out = 6)) +
  
  scale_x_continuous(name = "", 
                     breaks=seq(2,25, by = 2),
                     labels=month.name[c(3:12,1,2)]) +

  
  theme(legend.position = "none", 
        axis.text.x = element_text(size=14, angle=90), 
        axis.text.y = element_text(size=14),
        axis.title.y = element_text(size=24, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=24, margin = margin(t = 15, r = 0, b = 0, l = 0)))

# pop.acc



# positives

pos.acc=ggplot() + 
  
  geom_polygon(aes(x=c(1, 4, 4,1), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(16, 25, 25,16), y = c(-1.1,-1.1,3,3)), fill=alpha("#56B4E9",0.4))+
  geom_polygon(aes(x=c(4, 8, 8, 4), y = c(-1.1,-1.1,3,3)), fill=alpha("#FEE0B6",0.5))+
  geom_polygon(aes(x=c(8, 16, 16, 8), y = c(-1.1,-1.1,3,3)), fill=alpha("#DECBE4",0.5))+
  
  geom_line(data = dat.acc, aes(x = seq(2,24, by = 2), y = pos), color = 'red') +
  geom_point(data = dat.acc, aes(x = seq(2,24, by = 2,), y = pos, size=2), color = "red") +
  
  scale_y_continuous(
    name =expression(' '),
                     breaks=(seq(0,0.1, by = 0.0125)-attr(test3, 'scaled:center')) / attr(test3, 'scaled:scale'),
                     labels=seq(0,0.1, by = 0.0125)) +
  
  scale_x_continuous(name = "Month",
                     breaks=seq(2,25, by = 2),
                     labels=month.name[c(3:12,1,2)]) +
  
  theme(legend.position = "none", 
        axis.text.x = element_text(size=14, angle=90), 
        axis.text.y = element_text(size=14),
        axis.title.y = element_text(size=24, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=24, margin = margin(t = 15, r = 0, b = 0, l = 0)))


# pos.acc



# Put all plots in a single figure

ready=grid.arrange(grobs = list(precip.mor + theme(plot.margin = unit(c(0.2,0.2,0,0.2), "cm")),
                                precip.acc + theme(plot.margin = unit(c(0.2,0.2,0,0.2), "cm")),
                                pop.mor + theme(plot.margin = unit(c(0.2,0.2,0,0.2), "cm")),
                                pop.acc + theme(plot.margin = unit(c(0.2,0.2,0,0.2), "cm")),
                                pos.mor + theme(plot.margin = unit(c(0.2,0.2,0,0.2), "cm")),
                                pos.acc + theme(plot.margin = unit(c(0.2,0.2,0,0.2), "cm"))), 
                   nrow=3, ncol=2,
                   top = textGrob("   Morogoro, Tanzania                               Accra, Ghana",gp=gpar(fontsize=30,font=1)))


# ggsave(filename = "figure2.tiff", plot = ready, device = "png", width = 17*2, height = 40, units = "cm")


```

Overall proportion of positive feces 

```{r overallprop pos feces, message=FALSE, warning=FALSE}
mean(tz.dat$positive)
mean(gh.dat$positive)

```

Proportion of positive feces in the “lactation”, “weaning”, and the “rest of the year” periods

```{r prop pos feces, message=FALSE, warning=FALSE}
tz.dat%>%group_by(repro.per)%>%dplyr::summarize(prop.feces.per.period=mean(positive))
gh.dat%>%group_by(repro.per)%>%dplyr::summarize(prop.feces.per.period=mean(positive))

```

BLAST results for Acra

```{r blast results accra, message=FALSE, warning=FALSE}

blast.quan.gh=read.csv("ghana_blast_quan_results.csv")
blast.wata.gh=read.csv("ghana_blast_watanabe_results.csv")

blast.quan.gh=blast.quan.gh[blast.quan.gh$Organism=="Eidolon bat",]
blast.wata.gh=blast.wata.gh[blast.wata.gh$Organism=="Eidolon bat",]


blast.quan.gh=
blast.quan.gh %>% 
  # Within each grouping of A and B values.
  group_by( Query) %>% 
  # Sort rows in descending order by "value" column.
  arrange(desc( X..Pairwise.Identity)) %>% 
  # Pick the top 1 value
  slice(1) %>% 
  # Remember to ungroup in case you want to do further work without grouping.
  ungroup() %>%
  
  select(Maximum, X..Pairwise.Identity, Organism, Query, Accession)

blast.quan.gh$Query<-sapply(strsplit(blast.quan.gh$Query, "_"), function (x) x[1])


temp=gh.dat[gh.dat$positive==1,]$individual_id
temp=gsub(x =temp,  pattern = "GHAB0", replacement = "GHAB")
blast.quan.gh=blast.quan.gh[blast.quan.gh$Query%in%temp,]
# 
# #table(blast.quan.gh$Query)blast.wata.gh=
blast.quan.gh=
blast.quan.gh %>%
  # Within each grouping of A and B values.
  group_by( Query) %>%
  # Sort rows in descending order by "value" column.
  arrange(desc(X..Pairwise.Identity)) %>%
  # Pick the top 1 value
  slice(1) %>%
  ungroup()






blast.wata.gh=
blast.wata.gh %>% 
  # Within each grouping of A and B values.
  group_by( Query) %>% 
  # Sort rows in descending order by "value" column.
  arrange(desc(X..Pairwise.Identity)) %>% 
  # Pick the top 1 value
  slice(1) %>% 
  # Remember to ungroup in case you want to do further work without grouping.
  ungroup() %>%
  # select specific columns
  select(Maximum, X..Pairwise.Identity, Organism, Query, Accession)

blast.wata.gh$Query<-sapply(strsplit(blast.wata.gh$Query, "_"), function (x) x[1])

temp=gh.dat[gh.dat$positive==1,]$individual_id
temp=gsub(x =temp,  pattern = "GHAB0", replacement = "GHAB")
blast.wata.gh=blast.wata.gh[blast.wata.gh$Query%in%temp,]


blast.wata.gh=
blast.wata.gh %>% 
  # Within each grouping of A and B values.
  group_by(Query) %>% 
  # Sort rows in descending order by "value" column.
  arrange(desc(X..Pairwise.Identity)) %>% 
  # Pick the top 1 value
  slice(1)  %>%
  ungroup() 




blast.gh=
rbind(blast.quan.gh, blast.wata.gh)%>%
  group_by( Query) %>% 
  arrange(desc(X..Pairwise.Identity)) %>% 
  slice(1) 


summary(as.numeric(sub("%","",blast.gh$X..Pairwise.Identity)))

 # Find the organism of the matching sequence
Sys.setenv(ENTREZ_KEY='yourkey')

organism=rentrez::entrez_summary(db = "nuccore", id =blast.gh$Accession)
blast.gh$organism=unname(sapply(organism, function (x) x$organism, USE.NAMES = F))

```


BLAST results for Morogoro

```{r blast results morogoro, message=FALSE, warning=FALSE}


blast.quan.tz=read.csv("tz_blast_quan_results.csv")
blast.wata.tz=read.csv("tz_blast_watanabe_results.csv")

blast.quan.tz=blast.quan.tz[blast.quan.tz$Organism=="Eidolon bat",]
blast.wata.tz=blast.wata.tz[blast.wata.tz$Organism=="Eidolon bat",]


blast.quan.tz=
blast.quan.tz %>% 
  # Within each grouping of A and B values.
  group_by( Query) %>% 
  # Sort rows in descending order by "value" column.
  arrange(desc( X..Pairwise.Identity)) %>% 
  # Pick the top 1 value
  slice(1) %>% 
  # Remember to ungroup in case you want to do further work without grouping.
  ungroup() %>%
  
  select(Maximum, X..Pairwise.Identity, Organism, Query, Accession)

blast.quan.tz$Query<-sapply(strsplit(blast.quan.tz$Query, "_"), function (x) x[1])


temp=tz.dat[tz.dat$positive==1,]$individual_id
blast.quan.tz=blast.quan.tz[blast.quan.tz$Query%in%temp,]

#table(blast.quan.gh$Query)blast.wata.gh=
blast.quan.tz=
blast.quan.tz %>% 
  # Within each grouping of A and B values.
  group_by( Query) %>% 
  # Sort rows in descending order by "value" column.
  arrange(desc(X..Pairwise.Identity)) %>% 
  # Pick the top 1 value
  slice(1) %>%
  ungroup() 
  





blast.wata.tz=
blast.wata.tz %>% 
  # Within each grouping of A and B values.
  group_by( Query) %>% 
  # Sort rows in descending order by "value" column.
  arrange(desc(X..Pairwise.Identity)) %>% 
  # Pick the top 1 value
  slice(1) %>% 
  # Remember to ungroup in case you want to do further work without grouping.
  ungroup() %>%
  # select specific columns
  select(Maximum, X..Pairwise.Identity, Organism, Query, Accession)

blast.wata.tz$Query<-sapply(strsplit(blast.wata.tz$Query, "_"), function (x) x[1])

temp=tz.dat[tz.dat$positive==1,]$individual_id
blast.wata.tz=blast.wata.tz[blast.wata.tz$Query%in%temp,]


blast.wata.tz=
blast.wata.tz %>% 
  # Within each grouping of A and B values.
  group_by(Query) %>% 
  # Sort rows in descending order by "value" column.
  arrange(desc(X..Pairwise.Identity)) %>% 
  # Pick the top 1 value
  slice(1)  %>%
  ungroup() 




blast.tz=
rbind(blast.quan.tz, blast.wata.tz)%>%
  group_by( Query) %>% 
  arrange(desc(X..Pairwise.Identity)) %>% 
  slice(1) 


summary(as.numeric(sub("%","",blast.tz$X..Pairwise.Identity)))

 # Find the organism of the matching sequence
organism=rentrez::entrez_summary(db = "nuccore", id =blast.tz$Accession)
blast.tz$organism=unname(sapply(organism, function (x) x$organism, USE.NAMES = F))

```


Intercept versus Seasonal Model

```{r compare intercept seasonal models, message=FALSE, warning=FALSE, eval=T}

mod_intercept_stan=readRDS("model_intercept.RDS")
mod_sine_cos_stan=readRDS("model_sine_cosine.RDS")

# loo intercept model

log_lik = extract_log_lik(mod_intercept_stan, merge_chains = F, parameter_name = "log_lik")
rel_n_eff = relative_eff(exp(log_lik))
#m1.waic=waic(log_lik1) # 803.9
loo.intercept.only=loo(log_lik, r_eff =rel_n_eff ) # 803.9

# loo sine cosine model

log_lik = extract_log_lik(mod_sine_cos_stan, merge_chains = F, parameter_name = "log_lik")
rel_n_eff = relative_eff(exp(log_lik))
#m1.waic=waic(log_lik1) # 803.9
loo.sine.cosine=loo(log_lik, r_eff =rel_n_eff ) # 803.9

# Table
cbind(c("Intercept only", "Sine cosine"),
rbind(round(loo.intercept.only$estimates["looic",],3),
      round(loo.sine.cosine$estimates["looic",],3)))%>%
  kbl%>%
  kable_paper(full_width = F)
```

Table 1. The lower and upper endpoints of the 95% Highest Posterior Density Interval (HPDI) of the predicted 
monthly proportion of Eidolon helvum shedding coronaviruses in Morogoro, Tanzania produced by a sine-cosine 
model with a period of 12 months and a single annual maximum and minimum.

```{r Table 1, message=FALSE, warning=FALSE, eval=T}


#Table 1
samples.per.month=table(tz.dat$month)

month.index=as.numeric(factor(tz.dat$month, levels = month.name[c( 8:12,1:7)])) # August gets 1

Month.prop.pred=   purrr::map(sort(unique(month.index)),
                   function (x)
                   extract(mod_sine_cos_stan)$y_rep%>%
                   as_tibble()%>%
                   select(which(month.index==x))%>%
                   mutate(prop.pred = rowMeans(.))%>%
                   select(prop.pred))

Month.prop.pred.HPDI = purrr::map(Month.prop.pred,  
                                  function(x) round(HPDI(x, 0.95),3))



table1=data.frame(Month=month.name[c( 8:12,1:7)],
                  Min.HPDI=do.call(rbind,Month.prop.pred.HPDI)[,1],
                  Max.HPDI=do.call(rbind,Month.prop.pred.HPDI)[,2])
                  
 
table1%>%kbl()%>%kable_paper(full_width = F)



```

Posterior Probability Distributions (PProD) of the fixed effects model parameters. Table 2. Summary of the Posterior Probability Distributions of the “rest of the year” reproductive period parameter, the “weaning” reproductive period parameter, and the “lactation” reproductive period parameter. 

```{r Table 2, message=FALSE, warning=FALSE}

mod_fixed_stan = readRDS("model_repro_periods_only_fixed_effects.RDS")  


# Summary of the Posterior Probability Distributions of the sine-cosine logistic model parameters.

out=
cbind(
round(
  # summary mean sd of coeffiecients
  summary(mod_fixed_stan, pars = c("alpha",
                                        "beta_wean",
                                        "beta_lact"))$summary, 3)[,c("mean", "sd")],

#HPDI coefficients
do.call(rbind,
mclapply(c("alpha", "beta_wean", "beta_lact"), 
       function(x) 
       round(
        HPDI(
          unlist(c(extract(mod_fixed_stan)[x]), use.names = F), 0.95),
        3),
       mc.cores = 3)))

#Table
out %>%
  kbl() %>%
  kable_paper(full_width = F)


```

Odds ratios fixed effects model

```{r supporting OR fixed effects, message=FALSE, warning=FALSE}

OR.out=

rbind(

#weaning versus rest of the year


exp(
HPDI(c(extract(mod_fixed_stan)["beta_wean"][[1]]),
0.95)),

#lactation versus rest of the year


exp(
HPDI(c(extract(mod_fixed_stan)["beta_lact"][[1]]),
0.95)),

#weaning versus lactation

exp(
HPDI(c(extract(mod_fixed_stan)["beta_wean"][[1]] - extract(mod_fixed_stan)["beta_lact"][[1]]),
0.95)))

rownames(OR.out) = c("Weaning versus rest of the year",
                  "Lactation versus rest of the year",
                  "Weaning versus lactation")

#Table
OR.out %>%
  kbl() %>%
  kable_paper(full_width = F)

```


Figure 2. The Posterior Probability Distributions (darkblue) and the corresponding 95% Highest Posterior Density Interval (lightblue) of the odds ratio for coronavirus shedding by Eidolon helvum in the roost at Morogoro Tanzania during the “weaning” period versus the “rest of the year”, during the “weaning” period versus the “lactation” period, and during the “lactation” period versus the “rest of the year”. The vertical black line indicates a neutral odds ratio with value of one.

```{r Figure 2, message=FALSE, warning=FALSE}

dens.data=data.frame(wean.vs.rest=c(extract(mod_fixed_stan)["beta_wean"][[1]]),
                     lact.vs.rest=c(extract(mod_fixed_stan)["beta_lact"][[1]]),
                     wean.vs.lact=c(extract(mod_fixed_stan)["beta_wean"][[1]] - extract(mod_fixed_stan)["beta_lact"][[1]]))

# Wean versus rest of the year plot

dens <- density(exp(dens.data$wean.vs.rest))

data.all.dens=tibble(x = dens$x, y = dens$y)

data.hpdi <- tibble(x = dens$x, y = dens$y) %>% 
    mutate(variable = case_when(
      (x >= exp(hdi(dens.data$wean.vs.rest, .width = .95)[1]) &
       x <= exp(hdi(dens.data$wean.vs.rest, .width = .95)[2])) ~ "On",
      (x < exp(hdi(dens.data$wean.vs.rest, .width = .95)[1]) &
       x > exp(hdi(dens.data$wean.vs.rest, .width = .95)[2])) ~ "Off",
      TRUE ~ NA_character_))


wean.vs.rest=
  ggplot(data.all.dens, aes(x, y)) + geom_line() +
  geom_area(data = filter(data.all.dens), fill = 'lightblue') +
  geom_area(data = filter(data.hpdi, variable == 'On'), fill = 'cornflowerblue') + 
  theme(axis.text.x = element_text(size=14), 
        axis.text.y = element_text(size=14),
        axis.title.y = element_text(size=18, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=18, margin = margin(t = 15, r = 0, b = 0, l = 0)),
        plot.title = element_text(size=20, hjust = 0.5)) +
  scale_x_continuous(name ="Odds ratio", limits = c(-0.2,6)) +
  scale_y_continuous(name ="Density", limits = c(0,1.6)) +
  ggtitle("Weaning versus Rest of the year") +
  geom_vline(xintercept = 1) 



# Wean versus lactation

dens <- density(exp(dens.data$wean.vs.lact))

data.all.dens=tibble(x = dens$x, y = dens$y)

data.hpdi <- tibble(x = dens$x, y = dens$y) %>% 
    mutate(variable = case_when(
      (x >= exp(hdi(dens.data$wean.vs.lact, .width = .95)[1]) &
       x <= exp(hdi(dens.data$wean.vs.lact, .width = .95)[2])) ~ "On",
      (x < exp(hdi(dens.data$wean.vs.lact, .width = .95)[1]) &
       x > exp(hdi(dens.data$wean.vs.lact, .width = .95)[2])) ~ "Off",
      TRUE ~ NA_character_))


wean.vs.lact=
  ggplot(data.all.dens, aes(x, y)) + geom_line() +
  geom_area(data = filter(data.all.dens), fill = 'lightblue') +
  geom_area(data = filter(data.hpdi, variable == 'On'), fill = 'cornflowerblue') + 
  theme(axis.text.x = element_text(size=14), 
        axis.text.y = element_text(size=14, color="white"),
        axis.title.y = element_text(size=18, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=18, margin = margin(t = 15, r = 0, b = 0, l = 0)),
        plot.title = element_text(size=20, hjust = 0.5),
        axis.ticks.y = element_line(color="white")) +
  scale_x_continuous(name ="Odds ratio", limits = c(-0.2,6)) +
  scale_y_continuous(name =" ", limits = c(0,1.6)) +
  ggtitle("Weaning versus Lactation") +
  geom_vline(xintercept = 1) 



# Lactation versus Rest of the year

dens <- density(exp(dens.data$lact.vs.rest))

data.all.dens=tibble(x = dens$x, y = dens$y)

data.hpdi <- tibble(x = dens$x, y = dens$y) %>% 
    mutate(variable = case_when(
      (x >= exp(hdi(dens.data$lact.vs.rest, .width = .95)[1]) &
       x <= exp(hdi(dens.data$lact.vs.rest, .width = .95)[2])) ~ "On",
      (x < exp(hdi(dens.data$lact.vs.rest, .width = .95)[1]) &
       x > exp(hdi(dens.data$lact.vs.rest, .width = .95)[2])) ~ "Off",
      TRUE ~ NA_character_))


lact.vs.rest=
  ggplot(data.all.dens, aes(x, y)) + geom_line() +
  geom_area(data = filter(data.all.dens), fill = 'lightblue') +
  geom_area(data = filter(data.hpdi, variable == 'On'), fill = 'cornflowerblue') + 
  theme(axis.text.x = element_text(size=14), 
        axis.text.y = element_text(size=14, color="white"),
        axis.title.y = element_text(size=18, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=18, margin = margin(t = 15, r = 0, b = 0, l = 0)),
        plot.title = element_text(size=20, hjust = 0.5),
        axis.ticks.y = element_line(color="white")) +
  scale_x_continuous(name ="Odds ratio", limits = c(-0.2,6)) +
  scale_y_continuous(name =" ", limits = c(0,1.6)) +
  ggtitle("Lactation versus Rest of the year") +
  geom_vline(xintercept = 1) 



# Plot 


ready=grid.arrange(grobs = list(wean.vs.rest, #+ theme(plot.margin = unit(c(0,0,0,0), "cm")),
                                wean.vs.lact, #+ theme(plot.margin = unit(c(0,0.3,0,1.5), "cm")),
                                lact.vs.rest), #+ theme(plot.margin = unit(c(0,0.4,0,1.4), "cm"))), nrow=1)
                                nrow=1)

 # ggsave(filename = "figure3.tiff", plot = ready, device = "tiff", width = 40, height = 12, units = "cm")

ready

```

# DISCUSSION

Percentage of Eidolon bat coronavirus in E. helvum versus other coronaviruses detected in this species


```{r Eidolon bat coronavirus/Kenya/KY24/2006 detection in E. helvum, message=FALSE, warning=FALSE}

# Search in the nucleotide database of the national library of medicine
#((((deltacoronavirus) NOT human) NOT homo) NOT sapiens) 
#((((gammacoronavirus) NOT human) NOT homo) NOT sapiens) 
#((((betacoronavirus) NOT human) NOT homo) NOT sapiens) 
#((((alphacoronavirus) NOT human) NOT homo) NOT sapiens) 
#((((coronavirus) NOT human) NOT homo) NOT sapiens) 

gb.all=readRDS( "/Users/DMontecino/Documents/CoV_Wildlife/Data/COV_data_Feb_12_2022.RDS")

gb.all$CoV=NA

gb.all[is.na(gb.all$species),]$species="Unknown"



# General results


#total number of records 

nrow(gb.all)
nrow(gb.all[gb.all$species!="Unknown",])

#
sort(table(gb.all[gb.all$species!="Unknown",]$species, useNA = "always"))





#numbr of unique classes
gb.all[which(gb.all$class=="Bivalvia"),]$host="Chaerephon plicatus"
gb.all[which(gb.all$class=="Bivalvia"),]$species="Chaerephon plicatus"
gb.all[which(gb.all$class=="Bivalvia"),]$genus="Chaerephon"
gb.all[which(gb.all$class=="Bivalvia"),]$family="Molossidae"
gb.all[which(gb.all$class=="Bivalvia"),]$order="Chiroptera"
gb.all[which(gb.all$class=="Bivalvia"),]$class="Mammalia"


# unique(gb.all[gb.all$country=="Ghana" & gb.all$host=="Eidolon helvum",]$Description)
# unique(gb.all[gb.all$country=="Ghana" & gb.all$host=="Eidolon helvum",]$Accession)
# unique(gb.all[gb.all$country=="Ghana" & gb.all$host=="Eidolon helvum",]$Organism)
# 
# sort(unique(gb.all[gb.all$country=="Tanzania" & gb.all$host=="Eidolon helvum" & grepl(pattern = "TZAB", gb.all$Description),]$Description))
# unique(gb.all[gb.all$country=="Tanzania" & gb.all$host=="Eidolon helvum" & grepl(pattern = "TZAB", gb.all$Description),]$Accession)
# unique(gb.all[gb.all$country=="Tanzania" & gb.all$host=="Eidolon helvum" & grepl(pattern = "TZAB", gb.all$Description),]$Organism)


gb.all[ gb.all$host=="Eidolon helvum" & !grepl(pattern = "TZAB|GHAB", gb.all$Description),]$Organism

prop.table(table(gb.all[ gb.all$host=="Eidolon helvum" & !grepl(pattern = "TZAB|GHAB", gb.all$Description),]$Organism))*100
```

Other species where Eidolon bat coronavirus/Kenya/KY24/2006 has been reported

```{r Eidolon bat coronavirus/Kenya/KY24/2006 species detection in E. helvum, message=FALSE, warning=FALSE}

sort(unique(gb.all[ gb.all$Organism=="Eidolon bat coronavirus/Kenya/KY24/2006",]$host))


```

## Supporting Information 1

Summary rhat's

```{r supporting information 1 rhats, message=FALSE, warning=FALSE}
# rhat 
mod_hierarchical_stan<-readRDS("hierarchical_model.RDS")  



summ.rhat=lapply(
list(mod_intercept_stan, mod_sine_cos_stan, mod_fixed_stan, mod_hierarchical_stan), 
function(x)
summary(summary(x)$summary[,"Rhat"]))

summ.rhat=round(do.call(rbind,summ.rhat),4)

summ.rhat=
  cbind(c("Intercept model", 
            "Sine cosine model",
            "Fixed effects model",
              "Hierarchical model"),
  summ.rhat)


summ.rhat %>%
  kbl() %>%
  kable_paper(full_width = F)

```

Summary number of effective samples

```{r supporting information 1 num effective samples, message=FALSE, warning=FALSE}

summ.neff=lapply(
list(mod_intercept_stan, mod_sine_cos_stan, mod_fixed_stan, mod_hierarchical_stan), 
function(x)
summary(summary(x)$summary[,"n_eff"]))

summ.neff=round(do.call(rbind,summ.neff),4)

summ.neff=
  cbind(c("Intercept model", 
            "Sine cosine model",
            "Fixed effects model",
              "Hierarchical model"),
  summ.neff)


summ.neff %>%
  kbl() %>%
  kable_paper(full_width = F)

```


Summary of the Posterior Probability Distributions of the sine-cosine logistic model parameters.

```{r supporting information 1 sine - cosine model 1, message=FALSE, warning=FALSE}


# Summary of the Posterior Probability Distributions of the sine-cosine logistic model parameters.

out=
cbind(
round(
  # summary mean sd of coeffiecients
  summary(mod_sine_cos_stan, pars = c("alpha",
                                        "beta_sine",
                                        "beta_cosine"))$summary, 3)[,c("mean", "sd")],

#HPDI coefficients
do.call(rbind,
mclapply(c("alpha", "beta_sine", "beta_cosine"), 
       function(x) 
       round(
        HPDI(
          unlist(c(extract(mod_sine_cos_stan)[x]), use.names = F), 0.95),
        3),
       mc.cores = 3)))

#Table
out %>%
  kbl() %>%
  kable_paper(full_width = F)

```

Traceplots

```{r supporting information 1 sine - cosine model 2, message=FALSE, warning=FALSE}

#mod_sine_cos_stan<-readRDS( "model_sine_cosine.RDS")


# jpeg("Traceplots_sine_cosine_model_alpha.jpeg", width = 1100, height = 1000)
traceplot(mod_sine_cos_stan, pars="alpha") +
  
  theme(legend.title = element_text(size = 24),
        legend.text = element_text( size = 20), 
        axis.text.x = element_text(size=20), 
        axis.text.y = element_text(size=20),
        axis.title.y = element_text(size=34, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=34, margin = margin(t = 15, r = 0, b = 0, l = 0)))
# dev.off()


# jpeg("Traceplots_sine_cosine_model_beta_sine.jpeg", width = 1100, height = 1000)
traceplot(mod_sine_cos_stan, pars="beta_sine") +
  
  theme(legend.title = element_text(size = 24),
        legend.text = element_text( size = 20), 
        axis.text.x = element_text(size=20), 
        axis.text.y = element_text(size=20),
        axis.title.y = element_text(size=34, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=34, margin = margin(t = 15, r = 0, b = 0, l = 0)))
# dev.off()


# jpeg("Traceplots_sine_cosine_model_beta_cosine.jpeg", width = 1100, height = 1000)
traceplot(mod_sine_cos_stan, pars="beta_cosine") +
  
  theme(legend.title = element_text(size = 24),
        legend.text = element_text( size = 20), 
        axis.text.x = element_text(size=20), 
        axis.text.y = element_text(size=20),
        axis.title.y = element_text(size=34, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=34, margin = margin(t = 15, r = 0, b = 0, l = 0)))
# dev.off()
```

Fixed effects only

```{r supporting information 1 fixed effects model, message=FALSE, warning=FALSE}

# jpeg("Traceplots_repro_periods_model_alpha.jpeg", width = 1100, height = 1000)
traceplot(mod_fixed_stan, pars="alpha") +
  
  theme(legend.title = element_text(size = 24),
        legend.text = element_text( size = 20), 
        axis.text.x = element_text(size=20), 
        axis.text.y = element_text(size=20),
        axis.title.y = element_text(size=34, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=34, margin = margin(t = 15, r = 0, b = 0, l = 0)))
# dev.off()


# jpeg("Traceplots_repro_periods_model_beta_wean.jpeg", width = 1100, height = 1000)
traceplot(mod_fixed_stan, pars="beta_wean") +
  
  theme(legend.title = element_text(size = 24),
        legend.text = element_text( size = 20), 
        axis.text.x = element_text(size=20), 
        axis.text.y = element_text(size=20),
        axis.title.y = element_text(size=34, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=34, margin = margin(t = 15, r = 0, b = 0, l = 0)))
# dev.off()



# jpeg("Traceplots_repro_periods_model_beta_lact.jpeg", width = 1100, height = 1000)
traceplot(mod_fixed_stan, pars="beta_lact") +
  
  theme(legend.title = element_text(size = 24),
        legend.text = element_text( size = 20), 
        axis.text.x = element_text(size=20), 
        axis.text.y = element_text(size=20),
        axis.title.y = element_text(size=34, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=34, margin = margin(t = 15, r = 0, b = 0, l = 0)))
# dev.off()

```

Hierarchical model

```{r supporting information 1 hierarachical effects model, message=FALSE, warning=FALSE}



# repro periods traceplots

# jpeg("traceplots_hierarchical_repro_periods.jpeg", width = 1500, height = 500)
par(mfrow=c(3,1))

traceplot(mod_hierarchical_stan, pars=paste0("beta_repro[", c(1:3), "]")) +
  
  theme(legend.title = element_text(size = 24), #axis.text.x.top = element_text(size = 24),
        legend.text = element_text(size = 20), 
        axis.text.x = element_text(size=16), 
        axis.text.y = element_text(size=16),
        axis.title.y = element_text(size=34, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.title.x = element_text(size=34, margin = margin(t = 10, r = 0, b = 0, l = 0)))

# dev.off()





#month traceplots

# jpeg("traceplots_hierarchical_months.jpeg", width = 1500, height = 1000)
par(mfrow=c(4,3))

traceplot(mod_hierarchical_stan, pars=paste0("beta_month[", c(1:12), "]")) +
  
  theme(legend.title = element_text(size = 24),
        legend.text = element_text(size = 20), 
        axis.text.x = element_text(size=16), 
        axis.text.y = element_text(size=16),
        axis.title.y = element_text(size=34, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.title.x = element_text(size=34, margin = margin(t = 10, r = 0, b = 0, l = 0)),
        plot.title = element_text(size=30))

# dev.off()




# sigma repro traceplots

# jpeg("traceplots_hierarchical_sigma_repro.jpeg", width = 1500, height = 1000)
par(mfrow=c(1,1))

traceplot(mod_hierarchical_stan, pars="sigma_repro") +
  
  theme(legend.title = element_text(size = 30),
        legend.text = element_text( size = 24), 
        axis.text.x = element_text(size=25), 
        axis.text.y = element_text(size=25),
        axis.title.y = element_text(size=34, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=34, margin = margin(t = 15, r = 0, b = 0, l = 0)))

# dev.off()




# sigma month traceplots

# jpeg("traceplots_hierarchial_sigma_month.jpeg", width = 1500, height = 1000)
par(mfrow=c(1,1))

traceplot(mod_hierarchical_stan, pars="sigma_month") +
  
  theme(legend.title = element_text(size = 30),
        legend.text = element_text( size = 24), 
        axis.text.x = element_text(size=25), 
        axis.text.y = element_text(size=25),
        axis.title.y = element_text(size=34, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=34, margin = margin(t = 15, r = 0, b = 0, l = 0)))

# dev.off()

```


##Supporting Information 2

Summary of the Posterior Probability Distributions of the fixed effects model parameters.

```{r supporting information 2 fixed effects model, message=FALSE, warning=FALSE}


# Summary of the Posterior Probability Distributions of the sine-cosine logistic model parameters.

out=
cbind(
round(
  # summary mean sd of coeffiecients
  summary(mod_fixed_stan, pars = c("alpha",
                                        "beta_wean",
                                        "beta_lact"))$summary, 3)[,c("mean", "sd")],

#HPDI coefficients
do.call(rbind,
mclapply(c("alpha", "beta_wean", "beta_lact"), 
       function(x) 
       round(
        HPDI(
          unlist(c(extract(mod_fixed_stan)[x]), use.names = F), 0.95),
        3),
       mc.cores = 3)))

#Table
out %>%
  kbl() %>%
  kable_paper(full_width = F)

```


Summary of the Posterior Probability Distributions of the hierarchical model parameters.

```{r supporting information 3.1,,message=FALSE, warning=FALSE}

# Summary of the Posterior Probability Distributions of the hierarchical model parameters.

out=
cbind(

round(
  # summary mean sd of coeffiecients
  summary(mod_hierarchical_stan, pars = c("beta_repro",
                                        "beta_month",
                                        "sigma_repro",
                                        "sigma_month"))$summary, 3)[,c("mean", "sd")]

,

#HPDI coefficients

rbind(
do.call(rbind,
mclapply(c("beta_repro", "beta_month"),
function(x)         
t(round(
  apply(extract(mod_hierarchical_stan)[x][[1]], 
        MARGIN = 2, 
        function(y) HPDI(y, 0.95)),3)), mc.cores = 3)),

round(HPDI(c(extract(mod_hierarchical_stan)["sigma_repro"][[1]]), 0.95),3),

round(HPDI(c(extract(mod_hierarchical_stan)["sigma_month"][[1]]), 0.95),3)))


# Table hierarchical coefficients
out %>%
  kbl() %>%
  kable_paper(full_width = F)


```


Posterior probability distributions of the odds ratios of coronavirus shedding in E. helvum when comparing month m within reproductive period r and month m’ within reproductive period r’. The darkest blue shows the density within the 95% Highest Posterior Density Interval. The vertical lines show the odds ratio with value 1.


```{r supporting information 3.2,,message=FALSE, warning=FALSE}


# Posterior probability distributions of the odds ratios of coronavirus shedding in E. helvum when comparing month m within reproductive period r and month m’ within reproductive period r’. The darkest blue shows the density within the 95% Highest Posterior Density Interval. The vertical lines show the odds ratio with value 1.


temp=rstan::extract(mod_hierarchical_stan)



# Weaning versus rest of the year

wean.vs.rest.year=lapply(c(9,10), function(x) exp(temp$beta_0jk[,x]-temp$beta_0jk[,1]))

ap.vs.au=wean.vs.rest.year[[1]]

may.vs.au=wean.vs.rest.year[[2]]



# Lactation versus rest of the year

lact.versus.rest.year=lapply(c(6,7), function(x) exp(temp$beta_0jk[,x]-temp$beta_0jk[,1]))

jan.vs.au=lact.versus.rest.year[[1]]

feb.vs.au=lact.versus.rest.year[[2]]



# Weaning versus lactation

wean.versus.lact=lapply(c(9,10), function(x) exp(temp$beta_0jk[,x]-temp$beta_0jk[,7]))

ap.vs.feb=wean.versus.lact[[1]]

may.vs.feb=wean.versus.lact[[2]]




#Plot

densities=list(ap.vs.au, may.vs.au, jan.vs.au, feb.vs.au, ap.vs.feb, may.vs.feb)

titles=c("April vs Aug\nWeaning vs Rest of the year", "May vs Aug\nWeaning vs Rest of the year",
         "Jan vs Aug\nLactation vs Rest of the year", "Feb vs Aug\nLactation vs Rest of the year",
         "April vs Feb\nWeaning vs Lactation", "May vs Feb\nWeaning vs Lactation")

out=vector(mode = "list", length = length(titles))

for(i in 1:length(titles)){

g <- ggplot()
g <- g + geom_density(data=data.frame(value=densities[[i]]), aes(value), color=NA, fill="lightblue")

#building the data for the HPDI density
pr <- ggplot_build(g)

# subset the values within the HPDI
test=pr$data[[1]] %>% filter(x > hdi(densities[[i]], .width = .95)[1] & 
                             x < hdi(densities[[i]], .width = .95)[2])

# Plots
g =g + geom_ribbon(data = test, aes(x=x, ymin=0, ymax=y), inherit.aes=F,  fill="blue", alpha=0.3) +
  
  theme(axis.text.x = element_text(size=14), 
        axis.text.y = element_text(size=14),
        axis.title.y = element_text(size=18, margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(size=18, margin = margin(t = 15, r = 0, b = 0, l = 0)),
        plot.title = element_text(size=20, hjust = 0.5)) +
  scale_x_continuous(name =" ", limits = c(-0.2,10)) +
  scale_y_continuous(name =" ", limits = c(0,1.6)) +
  ggtitle(titles[i]) +
  geom_vline(xintercept = 1, lwd=2) +
  theme(plot.margin = unit(c(0,0.1,0,0), "cm"))

out[[i]]<-g}

for(i in c(1,3,5)){
  out[[i]]=out[[i]] + scale_y_continuous(name ="Density", limits = c(0,1.6))}

for(i in c(5,6)){
  out[[i]]=out[[i]] + scale_x_continuous(name ="Odds Ratio", limits = c(-0.2,10))}


ready=grid.arrange(grobs = out, nrow=3, ncol=2)


# ggsave(filename = "ods_ratios_hierarchical_model.png", 
#        plot = ready, 
#        device = "png", 
#        width = (40/3)*2, 
#        height = 12*3, 
#        units = "cm")
```

Posterior probability distributions of the reproductive period-level standard deviation (σR; pale red) and the month-level standard deviation (σM; pale green).

```{r supporting information 3.3,,message=FALSE, warning=FALSE}


# Posterior probability distributions of the reproductive period-level standard deviation (σR; pale red) and the month-level standard deviation (σM; pale green).

##plot

sigma.repro=temp$sigma_repro

sigma.month=temp$sigma_month

x <- data.frame(reproductive_period=sigma.repro, month=sigma.month)

data<- melt(x)

ready=ggplot(data,aes(x=value, fill=variable)) + 
  geom_density(alpha=0.25) +
  scale_x_continuous(name ="Standard deviation", limits = c(-0.2,12))


# ggsave(filename = "standard_deviations_hierarchical_model.png", 
#        plot = ready, 
#        device = "png", 
#        width = 25, 
#        height = 12, 
#        units = "cm")

```

##Supporting Information 3

```{r supporting information 3,,message=FALSE, warning=FALSE}


# Spearman
SpearmanRho(tapply(tz.dat$pos, list(tz.dat$month), mean),
            tapply(tz.dat$precip.month, list(tz.dat$month), unique),
            conf.level = 0.95)

plot( tapply(tz.dat$precip.month, list(tz.dat$month), unique), 
      tapply(tz.dat$pos, list(tz.dat$month), mean),
      main="Morogoro, lag 0",
      ylab="Monthly CoV detection",
      xlab="Precipitation")


#Kendall
KendallTauA(tapply(tz.dat$pos, list(tz.dat$month), mean),
            tapply(tz.dat$pop, list(tz.dat$month), unique),
            conf.level = 0.95)

# at the same time in Ghana

SpearmanRho(tapply(gh.dat$pos, list(gh.dat$month), mean),
            tapply(gh.dat$precip.month, list(gh.dat$month), unique),
            conf.level = 0.95)

plot(tapply(gh.dat$precip.month, list(gh.dat$month), unique),
     tapply(gh.dat$pos, list(gh.dat$month), mean),
     main="Accra, lag 0",
     ylab="Monthly CoV detection",
     xlab="Precipitation")


#Kendall
KendallTauA(tapply(gh.dat$pos, list(gh.dat$month), mean),
            tapply(gh.dat$precip.month, list(gh.dat$month), unique),
            conf.level = 0.95)


# between CoV detection and precipitation one month before  in Tanzania


# Spearman
SpearmanRho(tapply(tz.dat$pos, list(tz.dat$month), mean)[2:12],
            tapply(tz.dat$precip.month, list(tz.dat$month), unique)[1:11],
            conf.level = 0.95)

plot(tapply(tz.dat$precip.month, list(tz.dat$month), unique)[1:11],
     tapply(tz.dat$pos, list(tz.dat$month), mean)[2:12], 
     main="Morogoro, lag 1",
     ylab="Monthly CoV detection",
     xlab="Precipitation")


#Kendall
KendallTauA(tapply(tz.dat$pos, list(tz.dat$month), mean)[2:12],
            tapply(tz.dat$precip.month, list(tz.dat$month), unique)[1:11],
            conf.level = 0.95)


# between CoV detection and precipitation one month before  in Ghana

SpearmanRho(tapply(gh.dat$pos, list(gh.dat$month), mean)[2:12],
            tapply(gh.dat$precip.month, list(gh.dat$month), unique)[1:11],
            conf.level = 0.95)

plot(tapply(gh.dat$precip.month, list(gh.dat$month), unique)[1:11],
     tapply(gh.dat$pos, list(gh.dat$month), mean)[2:12],
     main="Accra, lag 1",
     ylab="Monthly CoV detection",
     xlab="Precipitation")


#Kendall
KendallTauA(tapply(gh.dat$pos, list(gh.dat$month), mean)[2:12],
            tapply(gh.dat$precip.month, list(gh.dat$month), unique)[1:11],
            conf.level = 0.95)


# between CoV detection and precipitation two months before in Tanzania

# Spearman
SpearmanRho(tapply(tz.dat$pos, list(tz.dat$month), mean)[3:12],
            tapply(tz.dat$precip.month, list(tz.dat$month), unique)[1:10],
            conf.level = 0.95)

plot(tapply(tz.dat$precip.month, list(tz.dat$month), unique)[1:10],
     tapply(tz.dat$pos, list(tz.dat$month), mean)[3:12],
     main="Morogoro, lag 2",
     ylab="Monthly CoV detection",
     xlab="Precipitation")
     


#Kendall
KendallTauA(tapply(tz.dat$pos, list(tz.dat$month), mean)[3:12],
            tapply(tz.dat$pop, list(tz.dat$month), unique)[1:10],
            conf.level = 0.95)


# between CoV detection and precipitation two months before in Ghana

SpearmanRho(tapply(gh.dat$pos, list(gh.dat$month), mean)[3:12],
            tapply(gh.dat$precip.month, list(gh.dat$month), unique)[1:10],
            conf.level = 0.95)

plot( tapply(gh.dat$precip.month, list(gh.dat$month), unique)[1:10],
      tapply(gh.dat$pos, list(gh.dat$month), mean)[3:12],
      main="Accra, lag 2",
      ylab="Monthly CoV detection",
      xlab="Precipitation")


#Kendall
KendallTauA(tapply(gh.dat$pos, list(gh.dat$month), mean)[3:12],
            tapply(gh.dat$precip.month, list(gh.dat$month), unique)[1:10],
            conf.level = 0.95)



SpearmanRho(tapply(gh.dat$pos, list(gh.dat$month), mean)[3:12],
            tapply(gh.dat$pop, list(gh.dat$month), unique)[1:10],
            conf.level = 0.95)

plot( tapply(gh.dat$pop, list(gh.dat$month), unique)[1:10],
      tapply(gh.dat$pos, list(gh.dat$month), mean)[3:12])



```

 